\cursedforest\ - A Random Forest Implementation for ``Big'' and ``Wide'' Data \\ Supplementary Information

### 2017-02-15 doc/publications/plos16/plos16.tex
changes to the conclusion

\cite{Donoho.and.Stodden.2006} investigate the behavior of a number of
regression approaches for variable selection in a small simulation.  They consider a regression problem with 
$\underset{n\times p}{X}\sim N(0,1)$ with $p$ fixed at 200 and $n$ variable.
They  set $\beta(1:k) \sim U(1,100)$ and $\beta((k+1):p) =0$, then  $y= X\beta + \epsilon$, $\epsilon \sim
N(0,\sqrt{16})$. They evaluate variable selection by the error measure 
$$\frac{\Vert\hat{\beta}-\beta\Vert_2}{\Vert\beta\Vert_2}$$

They consider a number of variable selection methods, including a false discovery rate criteria. This involves adding
the variable with the maximum $t$-value to the linear model if the $p$-value is less than 
$25(\text{number of  terms  currently  in  the  model})/(\text{total  number  of  variables})$.  
See Fig~\ref{figure:error_Stodden_FDR.png} for the error measure and
figure \ref{figure:rbo_Stodden_FDR.png} for the RBO, comparing the ranking (i.e. values) of $\hat{\beta}$ and $\beta$.
It apparent that there is a marked drop in the accuracy of the variable recovery in line with the prediction of the
Donoho-Tanner phase transition.

@@

We have seen no work on the question of such a phase transition in the case of classification problems. Still it may be
instructive to consider the implications of the phase transition for our work. As \cursedforest is designed for
extremely large numbers of variables it is likely to be operating in difficult regions of the figure where the ratio
$\delta = n/p$ is small. In the simulated example in section \ref{section:synthetic_data} the underdetermination $\delta = 0.002$ and the
sparsity $\rho =2\times 10^{-6}$

@@

We note that while a random forest is a long way from an all-subsets search, it is a limited combinatorial search of
the feature space. As such it may perform outside of the bounds of the phase transition. 
  combinatorial search.

@@

As \cursedforest is designed for extremely large numbers of variables it is likely to be operating in difficult regions of the figure.
We note that the examples we are considering lie on the extreme left of the plot, close to the origin. So there are
limits to what can be recovered in the presence of noise. However in the case of data that is both big and wide,
\cursedforest and other VariantSpark methods may provide a useful tool.

The existence of the  Donoho-Tanner phase transition is a salutary warning. There are likely to be limits, both
computational and logical, to the recovery of signals from noisy data. \cursedforest  is a contribution to addressing the
practical limits but the logical limits will still apply. However in the case of data that is both big and wide,
\cursedforest and other VariantSpark methods may provide a useful tool.

>>>>

\cite{Donoho.and.Stodden.2006} investigate the phase transition in a small simulation.  They consider a regression problem with 
$\underset{n\times p}{X}\sim N(0,1)$ with $p$ fixed at 200 and $n$ variable.
They  set $\beta(1:k) \sim U(1,100)$ and $\beta((k+1):p) =0$, then  $y= X\beta + \epsilon$, $\epsilon \sim
N(0,\sqrt{16})$. They evaluate variable selection by the error measure 
$$\frac{\Vert\hat{\beta}-\beta\Vert_2}{\Vert\beta\Vert_2}.$$

They consider a number of variable selection methods, including a false discovery rate criteria. This involves adding
the variable with the maximum $t$-value to the linear model if the $p$-value is less than 
$0.25(\text{number of  terms  currently  in  the  model})/(\text{total  number  of  variables})$.  
See Fig~\ref{figure:error_Stodden_FDR.png} for the error measure and
figure \ref{figure:rbo_Stodden_FDR.png} for the RBO, comparing the ranking (i.e. values) of $\hat{\beta}$ and $\beta$.
It apparent that there is a marked drop in the accuracy of the variable recovery in line with the prediction of the
Donoho-Tanner phase transition.

@@

We have seen no work on the question of such a phase transition in the case of classification problems. Still it may be
instructive to consider the implications of the phase transition for our work.

@@

We note that while a random forest is a long way from an all-subsets search, it is a limited search of
the feature space. As such it may perform outside of the bounds of the phase transition.

@@

As \cursedforest is designed for extremely large numbers of variables it is likely to be operating in difficult regions
of the figure where the ratio $\delta = n/p$ is small. In the simulated example in section \ref{section:synthetic_data}
the underdetermination $\delta = 0.002$ and the sparsity $\rho =2\times 10^{-6}.$ The existence of the Donoho-Tanner
phase transition is a salutary warning. There are likely to be limits, both computational and logical, to the recovery
of signals from noisy data. \cursedforest is a contribution to addressing the practical limits but the logical limits
will still apply. However in the case of data that is both big and wide, \cursedforest and other VariantSpark methods
may provide a useful tool.
